---
title: "Final Project Report - Practical Machine Learning Course"
author: "Mayank"
date: "22 April 2016"
output: html_document
---
#synopsis

"Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement ??? a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.


The following Libraries were used for this project, which you should install - if not done yet - and load on your working environment.

```{r, echo=TRUE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
```
Finally, load the same seed with the following line of code:
```{r, echo=TRUE}
set.seed(12345)
```

#Getting the data
The training data set can be found on the following URL:
```{r,echo=TRUE}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
```
The testing data set can be found on the following URL:

```{r,echo=TRUE}
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

```

#Partioning the training set into two
```{r,echo=TRUE}

inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]; myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)
```
#Cleaning the data
Cleaning Variables with too many NAs. For Variables that have more than a 60% threshold of NA's I'm going to leave them out:
```{r,echo=TRUE}
myTraining <- myTraining[c(-1)]
trainingV3 <- myTraining #creating another subset to iterate in loop
for(i in 1:length(myTraining)) { #for every column in the training dataset
        if( sum( is.na( myTraining[, i] ) ) /nrow(myTraining) >= .6 ) { #if n?? NAs > 60% of total observations
        for(j in 1:length(trainingV3)) {
            if( length( grep(names(myTraining[i]), names(trainingV3)[j]) ) ==1)  { #if the columns are the same:
                trainingV3 <- trainingV3[ , -j] #Remove that column
            }   
        } 
    }
}
#To check the new N?? of observations
dim(trainingV3)

myTraining <- trainingV3
rm(trainingV3)

```
Now let us do the exact same 3 transformations but for our myTesting and testing data sets.
```{r,echo=TRUE}
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -59]) #already with classe column removed
myTesting <- myTesting[clean1]
testing <- testing[clean2]

#To check the new N?? of observations
dim(myTesting)
#To check the new N?? of observations
dim(testing)
```
In order to ensure proper functioning of Decision Trees and especially RandomForest Algorithm with the Test data set (data set provided), we need to coerce the data into the same type.
```{r,echo=TRUE}
testing <- rbind(myTraining[2, -59] , testing) #note row 2 does not mean anything, this will be removed right.. now:
testing <- testing[-1,]

```

#Using ML algorithms for prediction: Decision Tree

```{r,echo=TRUE}
modFitA1 <- rpart(classe ~ ., data=myTraining, method="class")
fancyRpartPlot(modFitA1)
#predicting
predictionsA1 <- predict(modFitA1, myTesting, type = "class")
```
Using confusion Matrix to test results:
```{r,echo=TRUE}
confusionMatrix(predictionsA1, myTesting$classe)
```

#Using ML algorithms for prediction: Random Forests
```{r, echo=TRUE}
modFitB1 <- randomForest(classe ~. , data=myTraining)
#prediction
predictionsB1 <- predict(modFitB1, myTesting, type = "class")
```

Using confusion Matrix to test results:\
```{r,echo=TRUE}
confusionMatrix(predictionsB1, myTesting$classe)

```

Random Forests yielded better Results, as expected!


